---
title       : Kmeans versus Kernel Kmeans 
subtitle    : A Shiny Application Exploring Cluster Models
author      : Alex McBride
job         : 
framework   : io2012       # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : idea.css      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

---
## What is K-means?

K-means is a cluster analysis model used to cluster linear points in a data set

```{r, echo=FALSE, out.width='320px',out.height='200px'}
require(ggplot2) 
data(iris)
dataset <- iris[,1:2]
colnames(dataset) <- c("x","y")
Kclust <- kmeans(dataset, 2)
compute <- list(kmean.result = data.frame(dataset, cluster=as.factor(Kclust$cluster)))

ggplot(data=compute$kmean.result, aes(x=x, y=y, color=cluster)) +
                geom_point(size=3)
                
```

But how effective is it at clustering when the data is non-linear?

```{r, echo=FALSE, out.width='320px',out.height='200px'}
require("ggplot2")
set.seed(71)
data <- read.csv(file="self_test.data", head=TRUE, sep=" ")
names(data) <- c("X","Y")
p <- ggplot(data, aes(X, Y))
p + geom_point()
```

---
## K-means Cluster and Kernel K-means Cluster
```{r echo=FALSE, out.width='840px', out.height='230px'}
require("ggplot2")
require("gridExtra")
set.seed(71)
data <- read.csv(file="self_test.data", head=TRUE, sep=" ")
names(data) <- c("X","Y")
Kclust <- kmeans(data, center=2)
compute<- list(kmean.result = data.frame(data, cluster=as.factor(Kclust$cluster)))
plot1 <- ggplot(data=compute$kmean.result, aes(x=X, y=Y, color=cluster)) +
                geom_point(size=3)+
        ggtitle("K-means Clustering")

KKclust <- kkmeans(as.matrix(data), centers=2, kernel="rbfdot", 
                   kpar=list(sigma=1/(2*4^2)))
Kcompute <- list(kmean.result = data.frame(data, cluster=as.factor(KKclust@.Data)))
plot2 <- ggplot(data=Kcompute$kmean.result, aes(x=X, y=Y, color=cluster)) +
                geom_point(size=3)+
        ggtitle("Kernel K-means Clustering")
grid.arrange(plot1, plot2, ncol = 2)

```

## How does Kernel K-means differ?

Kernel k-means is an extension of the standard k-means algorithm that maps data points from input space to a higher dimensional feature space through a nonlinear transformation and minimizes the clustering error in feature space. But it does have its limitations....

---

## The Shiny Application

My shiny application, which you can run at [my Shiny App page,](https://lxdnz-254.shinyapps.io/ClusterAnalysis/ "Cluster ShinyApp") works by selecting the type of cluster modelling, (k-means or Kernel K-means) and number of clusters. From the inputs it produces a cluster plot showing the clustered dataset. It also runs a cross-validation test and gives results for Purity and NMI (Normalized Mutual Information).

In its current form, this application would be a useful teaching aid to show the difference between the two methods, and to highlight some of the drawbacks of each cluster model.
e.g. What happens when $\sigma$ (sigma) is set to high using the Kernel K-means algorithm?

The cross-validation test only gives a true result when the number of clusters is set to 2. For clusters larger than 2, the app generates a random test set based on the number of clusters selected and then produces results for NMI & Purity based on this random truth set. These results do not always reflect the exactness of the algorithms running.


---

## Further Developments 

This shiny app could become a very powerful tool, enabling quick cluster analysis on the fly.

It could be useful for presentations requiring various datasets, or to show different model results.
As a tool for data analysis, it could be used to visualise (quickly) some cluster model theories over a dataset, perhaps saving the user time by helping to rule out or select a model to optimise for their dataset.

In the future, I would like to incorporate into the Shiny App

 * Ability to choose datasets and variables
 * Select truth data sets
 * Select different types of cluster modelling
 * Change the style of plot output

To view the source code for the app or to collaborate visit my [Github Repo.](https://github.com/lxdnz254/ShinyApp)

